---
title: "Data_Processing_Final"
author: "Emine Cevik"
date: "`r Sys.Date()`"
output: html_document
---
```{r}
here::i_am("Code/Data_Processing_Final.Rmd")
```

```{r setup, include = TRUE, echo = FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r packages, warning = FALSE, echo = TRUE}
library(tidyverse)
library(data.table)
library(fpCompare)
library(openxlsx)
library(here)
library(agrmt)
options(scipen = 999)
```


# Basics

  - Read exp.txt file and create a data table 

```{r raw data loading, echo = TRUE}
exp_df <- read.table(here("Data", "exp.txt"))
exp_dt <- as.data.table(exp_df)

```

## Filtering

  - Time to filter the raw data 

  - As a first step, choose the entries with cert = TRUE and type = free
  --> free sampling paradigm where participants decide how much to sample and how to sample 
  --> Whether (at least) on of the options is a sure event

  - Secondly, eliminate the columns that we are not going to use 
  
    - outA3, outA4, outA5, outB3, outB4, outB5
    - probA3, probA4, probA5, probB3, probB4, probB5
    
```{r filter raw data, warning = FALSE, echo = TRUE}

filtered_exp <- exp_dt %>% filter(cert == TRUE, type =='free' )

filtered_exp[, c("outA3", "outA4", "outA5", "outB3", "outB4", "outB5", "probA3",
                 "probA4", "probA5", "probB3", "probB4", "probB5") := NULL]

```   

## Create identity attributions like participant

```{r new variables, echo = TRUE}

#identifier --> shows each decision experiment 
filtered_exp <- filtered_exp %>% 
  unite("identifier", c('id', 'subject', 'problem'), sep = "-", remove = FALSE)


#participant --> shows each decision maker 
filtered_exp <- filtered_exp %>% 
  unite("participant", c('id', 'subject'), sep = "-", remove = FALSE)


# study --> shows each different study 
filtered_exp <- as.data.table(filtered_exp)
filtered_exp[, study := substr(id, 1, 3) ]

```



## The descriptives of the data set 

```{r descriptives of the data set, echo = TRUE}

# Number of decisions made
filtered_exp[, uniqueN(identifier), by = dom]

# Number of participants 
filtered_exp[, uniqueN(participant), by = dom]

# Number of studies 
filtered_exp[, uniqueN(study), by = dom]

# Number of decisions made by  domains
#filtered_exp[, uniqueN(identifier), by = dom]

```

# Data processing

## Making data ready for analysis 
  - These changes are applied to implement the analysis easier 
  - When "choice" is equal to 0 --> A, 1 --> B
  - When "option" is equal to 0 --> A, 1 --> B
  - When "exval" is equal to 0 --> A, 1 --> B (also has NA)
  - When "und" is equal to 0 --> A, 1 --> B
  - When "cpt" is equal to 0 --> A, 1 --> B
  - When "risky" is equal to 0 --> A, 1 --> B (also has NA)
  - ev0 --> evA
  - ev1 --> evB
  
  Note: during our analysis, some of these changes are reversed

```{r data processing, warning = FALSE, echo = TRUE}

cleaned_exp <- as.data.table(filtered_exp)

cleaned_exp <- cleaned_exp %>% 
  mutate(choice = ifelse( choice == 0, 'A', 'B'))

cleaned_exp <- cleaned_exp %>% 
  mutate(option = ifelse( option == 0, 'A', 'B'))

cleaned_exp <- cleaned_exp %>% 
  mutate(und = ifelse( und == 0, 'A', 'B'))

cleaned_exp <- cleaned_exp %>% 
  mutate(cpt = ifelse( cpt == 0, 'A', 'B'))

cleaned_exp <- cleaned_exp %>% 
  mutate(exval = ifelse(exval == 0, "A", 
                        ifelse(exval == 1, "B", exval)))

cleaned_exp <- cleaned_exp %>% 
  mutate(risky = ifelse(risky == 0, "A", 
                        ifelse(risky == 1, "B", risky)))


# since values can be either A or B, created factor 
cleaned_exp$choice <- as.factor(cleaned_exp$choice)
cleaned_exp$option <- as.factor(cleaned_exp$option)
cleaned_exp$und <- as.factor(cleaned_exp$und)
cleaned_exp$cpt <- as.factor(cleaned_exp$cpt)
cleaned_exp$exval <- as.factor(cleaned_exp$exval)
cleaned_exp$risky <- as.factor(cleaned_exp$risky)

# rename ev0 and ev1 
cleaned_exp <- cleaned_exp %>% rename ("evA" = "ev0", 
                                       "evB" = "ev1")

```

## Observed properties 

  - seen_outcome (A1, A2, B1, B2): which outcome the participant saw at a trial
  - seen_option (same as option): : which option (A-B) the participant saw at a trial
  - seen_nout : number of distinct seen_outcome that the participant saw 
  - seen_all (boolean): shows if the participant saw all possible outcomes 
  - seen_A : how many time that the participant saw option A in the run
  - seen_B : how many time that the participant saw option B in the run
  - seen_A1 : how many time that the participant saw outcome A1 in the run
  - seen_A2 : how many time that the participant saw outcome A2 in the run
  - seen_B1 : how many time that the participant saw outcome B1 in the run
  - seen_B2 : how many time that the participant saw outcome B2 in the run
  - seen_probA1 = seen_A1 / seen_A
  - seen_probA2 = seen_A2 / seen_A
  - seen_probB1 = seen_B1 / seen_B
  - seen_probB2 = seen_B2 / seen_B
  - seen_evA = seen_probA1 * outA1 + seen_probA2 * outA2
  - seen_evB = seen_probB1 * outB1 + seen_probB2 * outB2
  - seen_covA = seen_stdA / seen_meanA
    - seen_meanA : for each trial, the average of the outcomes when 
    the seen_option is equal to A 
    - seen_stdA : for each trial, the standard deviation of the outcomes 
    when the seen_option is equal to B 
  - seen_covB = seen_stdB / seen_meanB
    - seen_meanB : for each trial, the average of the outcomes when 
    the seen_option is equal to A 
    - seen_stdB : for each trial, the sta
  - seen_risky 
  

```{r observed properties , echo = TRUE}

# seen_outcome
# note: it is important to use option condition because outcomes could be same in option A and B
cleaned_exp <- cleaned_exp %>% mutate(seen_outcome = case_when 
                                      ( option == 'A' & outcome == outA1 ~ 'A1',
                                        option == 'A' & outcome == outA2 ~ 'A2',
                                        option == 'B' & outcome == outB1 ~ 'B1',
                                        option == 'B' & outcome == outB2 ~ 'B2'))


# seen_option : even though it is exactly same with option column, I want to have it 
# for sanitary check later on
cleaned_exp <- cleaned_exp %>% mutate(seen_option = option)


# seen_nout : number of distinct seen_outcome that the participant saw 
cleaned_exp <- cleaned_exp %>% 
  group_by(identifier) %>% 
  mutate(seen_nout = n_distinct(seen_outcome))


# seen_all : number of distinct seen_outcome that the participant saw 

cleaned_exp <- as.data.table(cleaned_exp)
cleaned_exp[, seen_all := ifelse(nout == seen_nout, TRUE, ifelse(nout > seen_nout, FALSE, NA))]


# seen_A : how many time that the participant saw option A in the run
cleaned_exp <- cleaned_exp %>%  group_by(identifier) %>% 
  mutate(seen_A = sum(option == "A")) %>% 
  ungroup()


# seen_B : how many time that the participant saw option B in the run
cleaned_exp <- cleaned_exp %>%  group_by(identifier) %>% 
  mutate(seen_B = sum(option == "B")) %>% 
  ungroup()

# seen_A1 : how many time that the participant saw outcome A1 in the run
cleaned_exp <- cleaned_exp %>%  group_by(identifier) %>% 
  mutate(seen_A1 = sum(seen_outcome == "A1")) %>% 
  ungroup()


# seen_A2 : how many time that the participant saw outcome A2 in the run
cleaned_exp <- cleaned_exp %>%  group_by(identifier) %>% 
  mutate(seen_A2 = sum(seen_outcome == "A2")) %>% 
  ungroup()


# seen_B1 : how many time that the participant saw outcome B1 in the run
cleaned_exp <- cleaned_exp %>%  group_by(identifier) %>% 
  mutate(seen_B1 = sum(seen_outcome == "B1")) %>% 
  ungroup()


# seen_B2 : how many time that the participant saw outcome B2 in the run
cleaned_exp <- cleaned_exp %>%  group_by(identifier) %>% 
  mutate(seen_B2 = sum(seen_outcome == "B2")) %>% 
  ungroup()


# seen_probA1 = seen_A1/seen_A
cleaned_exp <- as.data.table(cleaned_exp)
cleaned_exp[, seen_probA1 := round(seen_A1 / seen_A, 3)]


# seen_probA2 = seen_A2/seen_A
cleaned_exp[, seen_probA2 := round(seen_A2 / seen_A, 3)]


# seen_probB1 = seen_B1/seen_B
cleaned_exp[, seen_probB1 := round(seen_B1 / seen_B, 3)]


# seen_probB2 = seen_B2/seen_B
cleaned_exp[, seen_probB2 := round(seen_B2 / seen_B, 3)]


# seen_evA = outA1*seen_probA1 + outA2*seen_probA2
cleaned_exp[, seen_evA := round(outA1*seen_probA1 + outA2*seen_probA2, 3)]


# seen_evB = outB1*seen_probB1 + outB2*seen_probB2
cleaned_exp[, seen_evB := round(outB1*seen_probB1 + outB2*seen_probB2, 3)]


# Convert "outcome" to numeric
cleaned_exp[, outcome := as.numeric(outcome)]

# In order to calculate coefficient of variation, we need to calculate mean and std 

# seen_meanA 
cleaned_exp <- cleaned_exp %>%
  group_by(identifier) %>%
  mutate(seen_meanA = mean(outcome[option == "A"], na.rm = TRUE)) %>%
  ungroup()

# seen_meanB
cleaned_exp <- cleaned_exp %>%
  group_by(identifier) %>%
  mutate(seen_meanB = mean(outcome[option == "B"], na.rm = TRUE)) %>%
  ungroup()


# seen_stdA
cleaned_exp <- cleaned_exp %>%
  group_by(identifier) %>%
  mutate(seen_stdA = sd(outcome[option == "A"], na.rm = TRUE)) %>%
  ungroup()

# seen_stdB
cleaned_exp <- cleaned_exp %>%
  group_by(identifier) %>%
  mutate(seen_stdB = sd(outcome[option == "B"], na.rm = TRUE)) %>%
  ungroup()

# seen_covA
cleaned_exp <- as.data.table(cleaned_exp)

cleaned_exp[, seen_covA := (seen_stdA / seen_meanA) , by = identifier]

# seen_covB
cleaned_exp[, seen_covB := (seen_stdB / seen_meanB) , by = identifier]

# seen_risky
cleaned_exp[, seen_risky := ifelse(is.na(seen_covA) & !is.na(seen_covB), "B",
                                 ifelse(!is.na(seen_covA) & is.na(seen_covB), "A", 
                                        ifelse(seen_covA > seen_covB, "A",
                                               ifelse(seen_covB > seen_covA, "B", NA)))
                                 )]

```

## Observed option with a higher expected return 
a winner of the summary decision strategy

```{r seen_exval , echo = TRUE}

cleaned_exp[, seen_exval := ifelse(seen_evA > seen_evB, "A",
                                   ifelse(seen_evA < seen_evB, "B", NA_character_))]

```

## Observed option as the winner of the round-wise decision strategy 

```{r seen winner, echo = FALSE}
cleaned_exp <- cleaned_exp %>% mutate(seen_outcomeA_higher_prob := case_when 
                                      (seen_probA1 > seen_probA2 ~ outA1,
                                       seen_probA1 < seen_probA2 ~ outA2))

cleaned_exp <- cleaned_exp %>% mutate(seen_outcomeB_higher_prob := case_when 
                                      ( seen_probB1 > seen_probB2 ~ outB1,
                                        seen_probB1 < seen_probB2 ~ outB2))

cleaned_exp[, seen_winner := ifelse( seen_outcomeA_higher_prob > seen_outcomeB_higher_prob, "A", "B") ]

# seen_winner will return NA in such cases:
# - seen_probA1 and seen_probA2 are 0.5 or seen_probB1 and seen_probB2 are 0.5
# - only one of the option was sampled (eg: never sampled from option B)

```

## Properties of the sampling process
  - sampling_total: total number of samples made before the decision
  - sampling_switch : shows whether the option sampled is same with the previous option sampled
  - sampling_switch_total : total number of switched occurred before the decision
  - sampling_switch_ratio = sampling_switch_total/(sampling_total - 1)

```{r sampling variables, echo = FALSE}

# sampling_total : number of trials by identifier 
cleaned_exp[, sampling_total := max(trial), by = identifier]


# sampling switch 
cleaned_exp[, sampling_switch := ifelse(trial == 1, "first",
                                         ifelse(option != lag(option), "yes", "no")), by = identifier]


# sampling_total_switch 
cleaned_exp[, sampling_total_switch := sum(sampling_switch == "yes"), by = identifier]


# sampling_switch_ratio (sampling_total_switch/(sampling_total-1))
cleaned_exp[, sampling_switch_ratio := round(sampling_total_switch/(sampling_total-1), 3), 
            by = identifier]

# --> high switch ratio means that sampling_total_switch and sampling_total are close 

```


# Another definition of roundwise winner
Finding the winner of the each small round and compare the number of winners that 
an option has 

## sampling round : uninterrepted sequence of options 
  
```{r sampling_round, echo = TRUE}

# create a temp data table to keep only variables required to calculate 
# sampling_round 

temp <- cleaned_exp[, c("identifier", "sampling_switch" )]

# Initialize "number of yes" to 0 for each identifier
temp$number_of_yes <- 0

# Update "number of yes" based on "sampling_switch" column
temp[, number_of_yes := cumsum(sampling_switch == "yes"), by = identifier]

# Initialize round column with NAs
temp$round <- NA

# Iterate through each row
current_round <- 1

for (i in 1:nrow(temp)) {

  if (temp$sampling_switch[i] == "first") {
    temp$round[i] <- 1
    current_round <- 1
  }

  if (i > 1 && temp$identifier[i] == temp$identifier[i - 1]) {
    if (temp$number_of_yes[i] == temp$number_of_yes[i - 1]) {
      temp$round[i] <- current_round
    } else {
      if (temp$number_of_yes[i] %% 2 != 0) {
        temp$round[i] <- current_round
      } else {
        current_round <- current_round + 1
        temp$round[i] <- current_round
      }
    }
  }
}

cleaned_exp[, sampling_round := temp$round]

cleaned_exp[, round := temp$round]
```

## Round_completed
round means uninterrepted sequence of options. so, some of the last rounds might be uncompleted.

```{r round completed}

# create column called round completed
cleaned_exp[, round_completed := ifelse(uniqueN(option) == 2, TRUE, FALSE), by = c("identifier", "round")]

```

## round winner: the option with the higher sampled average return for each round

  - round_seen_A & round_seen_B: how many times option A & B was seen during the run
  - round_seen_A1 & round_seen_B1 : how many times outcome A1 & B1 was seen during the run
  - round_seen_A2 & round_seen_B2 : how many times outcome A2 & B2 was seen during the run
  - round_seen_probA1 & round_seen_probB1 : round_seen_A1 / round_seen_A
  - round_seen_probA2 & round_seen_probB2 : round_seen_A2 / round_seen_A
  - round_seen_evA & round_seen_evB : round_seen_probA1 * outA1 + round_seen_probA2 * outA2
  - round_winner : comparison between round_seen_evA and round_seen_evB
  - roundwise_winner : most frequent round_winner
  
```{r required variables for round_winner}

# round_seen_A
# count option A in a the round 
cleaned_exp <- cleaned_exp %>%  group_by(identifier, round) %>% 
  mutate(round_seen_A = sum(option == "A")) %>% 
  ungroup()


# round_seen_B
# count option B in a the round
cleaned_exp <- cleaned_exp %>%  group_by(identifier, round) %>% 
  mutate(round_seen_B = sum(option == "B")) %>% 
  ungroup()



# round_seen_A1 (how many time that the participant saw outcome A1 in the round)
cleaned_exp <- cleaned_exp %>%  group_by(identifier, round) %>% 
  mutate(round_seen_A1 = sum(seen_outcome == "A1")) %>% 
  ungroup()


# round_seen_A2 (how many time that the participant saw outcome A2 in the round)
cleaned_exp <- cleaned_exp %>%  group_by(identifier, round) %>% 
  mutate(round_seen_A2 = sum(seen_outcome == "A2")) %>% 
  ungroup()


# round_seen_B1 (how many time that the participant saw outcome B1 in the round)
cleaned_exp <- cleaned_exp %>%  group_by(identifier, round) %>% 
  mutate(round_seen_B1 = sum(seen_outcome == "B1")) %>% 
  ungroup()


# round_seen_B2 (how many time that the participant saw outcome B2 in the run)
cleaned_exp <- cleaned_exp %>%  group_by(identifier, round) %>% 
  mutate(round_seen_B2 = sum(seen_outcome == "B2")) %>% 
  ungroup()


# round_seen_probA1 = round_seen_A1/round_seen_A
cleaned_exp <- as.data.table(cleaned_exp)
cleaned_exp[, round_seen_probA1 := round(round_seen_A1 / round_seen_A, 3)]


# round_seen_probA2 = round_seen_A2/round_seen_A
cleaned_exp[, round_seen_probA2 := round(round_seen_A2 / round_seen_A, 3)]


# round_seen_probB1 = round_seen_B1/round_seen_B
cleaned_exp[, round_seen_probB1 := round(round_seen_B1 / round_seen_B, 3)]


# round_seen_probB2 = round_seen_B2/round_seen_B
cleaned_exp[, round_seen_probB2 := round(round_seen_B2 / round_seen_B, 3)]


# round_seen_evA = outA1*round_seen_probA1 + outA2*round_seen_probA2
cleaned_exp[, round_seen_evA := round(outA1*round_seen_probA1 + outA2*round_seen_probA2, 3)]


# round_seen_evB = outB1*round_seen_probB1 + outB2*round_seen_probB2
cleaned_exp[, round_seen_evB := round(outB1*round_seen_probB1 + outB2*round_seen_probB2, 3)]


```


## round_winner : based on the comparison between round_seen_evA and round_seen_evB

When round is *completed*
  - round_seen_evA > round_seen_evB --> round_winner : A
  - round_seen_evB > round_seen_evA --> round_winner : B
  - round_seen_evA = round_seen_evA --> round_winner : no winner 

When round is *uncomplete* 
  - round_winner : undefined 
  
  
```{r round winner}
# The case we don't include uncompleted rounds to calculation
cleaned_exp[, round_winner := ifelse( round_completed == FALSE, "undefined", 
                                      ifelse(round_seen_evA > round_seen_evB, "A", 
                                             ifelse(round_seen_evA < round_seen_evB, "B", 
                                                    "no winner")))]

```

## roundwise winner 

Based on only *completed* rounds, compare the number of times A and B 
were winners of a round

Step 1
  - round_winner_A : how many times A was the round winner 
  - round_winner_B : how many times B was the round winner 

Step 2  
roundwise_winner 
  - round_winner_A > round_winner_B --> roundwise_winner : A 
  - round_winner_B > round_winner_A --> roundwise_winner : B   
  - round_winner_A = round_winner_B --> roundwise_winner : no winner  


```{r round winner A and round winner B}

temp2 <- cleaned_exp[, c("identifier", "trial", "sampling_round", "round", "round_winner")]

# Initialize round_winner_A and round_winner_B for each identifier
temp2$round_winner_A <- 0
temp2$round_winner_B <- 0


for(i in 1:nrow (temp2)) {
  if (i == 1)
  {
    total_A <- 0
    total_B <- 0
    
    if(temp2$round_winner[i] == "A") {
      total_A <- total_A + 1 
      temp2$round_winner_A[i] <- total_A
      temp2$round_winner_B[i] <- total_B
      
    } else if (temp2$round_winner[i] == "B") {
      total_B <- total_B + 1 
      temp2$round_winner_A[i] <- total_A
      temp2$round_winner_B[i] <- total_B
      
    } else {
      temp2$round_winner_A[i] <- total_A
      temp2$round_winner_B[i] <- total_B
    }
  } else {
    
    # check if we are still looking at the same identifier 
    if(temp2$identifier[i] == temp2$identifier[i-1]) {
      
      # check if we are still looking at the same round 
      if(temp2$round[i] == temp2$round[i-1]) {
        temp2$round_winner_A[i] <- temp2$round_winner_A[i-1]
        temp2$round_winner_B[i] <- temp2$round_winner_B[i-1]     
      } else {
      if(temp2$round_winner[i] == "A") {
        total_A <- total_A + 1 
      } else if (temp2$round_winner[i] == "B") {
        total_B <- total_B + 1 
      } else {
        total_A <- total_A + 0
        total_B <- total_B + 0
      }
        temp2$round_winner_A[i] <- total_A
        temp2$round_winner_B[i] <- total_B
    }
    } else {
      total_A <- 0
      total_B <- 0 
      
      if (temp2$round_winner[i] == "A") {
        total_A <- total_A + 1 
      } else if (temp2$round_winner[i] == "B") {
        total_B <- total_B + 1 
      } else {
        total_A <- total_A + 0
        total_B <- total_B + 0
      }
      
      temp2$round_winner_A[i] <- total_A
      temp2$round_winner_B[i] <- total_B
  }
  }
}
  
cleaned_exp$round_winner_A <- temp2$round_winner_A
cleaned_exp$round_winner_B <- temp2$round_winner_B

```

Step 2 : Roundwise winner 
```{r roundwise winner}

temp2 <- as.data.table(temp2)

temp2[, roundwise_winner := ifelse(max(round_winner_A) == max(round_winner_B), "no winner",
                                                    ifelse(max(round_winner_A) > max(round_winner_B), "A", "B")),
                     by = identifier]

cleaned_exp$roundwise_winner <- temp2$roundwise_winner
```

# Choice related properties

## Which decision strategy is chosen?
  - seen_winner_chosen: if seen_winner and choice are the same 
  - seen_exval_chosen: if seen_exval and choice are the same 
  - seen_roundwise_winner_chosen : if the seen_roundwise_winner and choice are the same 


```{r choice behavior}

cleaned_exp[, seen_exval_chosen := ifelse(seen_exval == choice, TRUE, FALSE), by = identifier]

cleaned_exp[, seen_winner_chosen := ifelse(seen_winner == choice, TRUE, FALSE), by = identifier]

cleaned_exp[, seen_roundwise_winner_chosen := ifelse(roundwise_winner == choice, TRUE, FALSE), by = identifier]

```


# Create an indicator variable for the analysis 
  - I1: A variable that shows if seen_exval and seen_winner are same
  - I2: A variable that shows if seen_exval and roundwise_winner are same

```{r indicator variable }

cleaned_exp[, I1 := ifelse(seen_exval == seen_winner, TRUE, FALSE) ]
cleaned_exp[, I2 := ifelse(seen_exval == roundwise_winner, TRUE, FALSE)]

```

# Sampling with completed rounds
```{r}
cleaned_exp[, sampling_uncompleted_round := ifelse(sum(round_completed == FALSE) > 0, TRUE, FALSE),
            by = identifier]
```


# Create a data set without trial details 
```{r}
subset_dt <- cleaned_exp[trial == 1 , ]
```

# Write this data set to csv 
```{r}
write.csv(subset_dt, file = here("Data", "final_analysis_ready_data.csv"), row.names = FALSE)
```


