---
title: "Descriptive_Analysis_final"
author: "Emine Cevik"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
here::i_am("Code/Descriptive_Analysis_Final.Rmd")
```

```{r setup, include = TRUE, echo = FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r, warning = FALSE, echo = TRUE}
library(tidyverse)
library(openxlsx)
library(data.table)
library(ggplot2)
library(here)
library(ggpubr)
library(patchwork)
```

# Read data set
```{r read data set, echo = FALSE}
#read the file 
main_dt <- as.data.table(read.csv(here("Data", "final_analysis_ready_data.csv")))
colnames(main_dt)
```

# Filter the decisions where participants didn't sample more than once 
```{r}
main_dt <- main_dt[main_dt[, sampling_total > 1]]
```

# Filter the decisions where participants didn't sample from both of the option
```{r}
cat("in ", main_dt[ seen_A == 0 | seen_B == 0, .N], " of the decisions, participants did not sample from both option\n")

main_dt <- main_dt[main_dt[, seen_A > 0 & seen_B > 0 ]]
```


# Separate data based on domain 
  - Since the results of ther analysis were similar over different domains, I decided to use only gain domain. 
  - However, all operations will be also done for loss and mixed domain at the end
```{r separate data}
gain_dt <- main_dt[dom == "Gain", ]
loss_dt <- main_dt[dom == "Loss", ]
mixed_dt <- main_dt[dom == "Mixed", ]
```

# Check descriptive of the gain data 
```{r, warning = FALSE, echo = TRUE}
# Number of different session (identifier) 
gain_dt[, .N]

# Total number of participants 
gain_dt[, uniqueN(participant)]

# Total number of trials 
gain_dt[, sum(sampling_total)]

# Average length of sampling sequence by identifier 
gain_dt[, mean(sampling_total)]

# Average identifier by participant 
# which shows the average number of problems that each participant solve 
gain_dt[, .N]/gain_dt[, uniqueN(participant)]

# Total number of studies
gain_dt[, uniqueN(study)]

```

# Hills and Hertwig sample error control

  - experiments where participants see all outcomes at least once
  
  - Did participants see all outcomes within the trial?

```{r all outcomes are seen}

total_seen_all <- gain_dt[seen_all == "TRUE", .N]
percentage_of_seen_all <- round(total_seen_all/gain_dt[, .N], 3)

print(paste0("Only in ", percentage_of_seen_all*100, "% of the run, participants saw all the outcomes"))

```

# Sampling total 1
One of the properties of the sampling paradigm is that participants have a full control over how they sample 
Previous findings foudn that people search frugally 

```{r descriptives of length of the sampling sequence}

cat("mean: ", mean(gain_dt$sampling_total), "\n")
cat("median: ", median(gain_dt$sampling_total), "\n")
cat("max: ", max(gain_dt$sampling_total), "\n")
cat("min: ", min(gain_dt$sampling_total), "\n")

```

# Sampling total 2
```{r histogram sampling total}

# Create a histogram with a bin size of  
# added median line to the graph

plot1 <- ggplot(gain_dt, aes(x = sampling_total)) + geom_histogram(binwidth = 3, fill="black", col="grey") + 
  geom_vline(xintercept = median(gain_dt$sampling_total), linetype = "dashed", color = "blue", size = 1) + 
  annotate("text", x = 80, y = 750, label="Median sample size is 15", angle = 0, size = 6) +
  labs(title = "Histogram of the length of the sampling sequences", 
       x = "Sampling size",y = "Frequency", size = 35) + xlim(c(0,268)) + theme_classic() + 
  theme(
  plot.title = element_text(color = "black", size = 16, face = "bold",hjust = 0.5),
  axis.title = element_text(size = 16)
) 


plot1

#
ggsave("plot1.png", plot1)
 
```

# Sampling frequency 
It is an important metric to be checked in detail. 

  - When the sampling frequency is high, it is piecewise sampling 
  - When the sampling frequency is low, it is comprehensive sampling strategy 
  
```{r descriptives of length of the sampling frequency}

cat("mean: ", mean(gain_dt$sampling_switch_ratio), "\n")
cat("median: ", median(gain_dt$sampling_switch_ratio), "\n")
cat("max: ", max(gain_dt$sampling_switch_ratio), "\n")
cat("min: ", min(gain_dt$sampling_switch_ratio), "\n")

```  
  

```{r sampling frequency}
# cleaned_exp[, sampling_switch_ratio := round(sampling_total_switch/(sampling_total-1), 3), by = identifier]

# Based on the definition, if the sampling total is equal to 1, the switch ratio will be NA 

# Create a histogram 
# added median line to the graph


plot2 <- ggplot(gain_dt, aes(x = sampling_switch_ratio)) + geom_histogram( fill="black", col="grey") + 
  geom_vline(xintercept = median(gain_dt$sampling_switch_ratio, na.rm = TRUE), linetype = "dashed", color = "blue", size = 1) + 
  labs(title = "Histogram of switching frequency",
       x = "Switching frequency",
       y = "Number of decisions") + annotate("text", x = 0.62, y = 1000, label="Median switching frequency is 0.29", angle = 0, size = 6) + theme_classic() + 
  theme(
  plot.title = element_text(color = "black", size = 16, face = "bold",hjust = 0.5),
  axis.title = element_text(size = 16)
) 

plot2

ggsave("plot2.png", plot2)


```

It is interesting to see high number of cases where switching ratio is equal to 1. 

(When it is closer to 1, it shows frequent switching behavior)

Also it seems like there are clearly two different groups which is what we can pursue and ask questions about them 

--> Bimodel distribution

# How about the relationship between sampling frequency vs sampling total 

```{r sampling frequency vs sampling total}

cols <- c("Mean sample size by switcing frequency"="blue","Instance"="grey")

# find the mean sampling_switch_ratio by sampling_total 
gain_dt[, sampling_switch_ratio_mean := round(mean(sampling_switch_ratio),2), by = sampling_total]

# find the mean sampling_total by sampling_switch_ratio
gain_dt[, sampling_total_mean := round(mean(sampling_total),2), by = sampling_switch_ratio]

plot3 <- ggplot(gain_dt) + geom_point( aes( x = sampling_total, y = sampling_switch_ratio), color = "gray", fill = "gray") +
  geom_point(aes( x = sampling_total, y = sampling_switch_ratio_mean) ,color = "blue", fill = "blue") +
  labs(title = "The relationship between sample size and switching frequency", 
       x = "Sample size", y = "Switching frequency") +  scale_fill_manual(values = c("blue", "grey")) +theme_classic() + 
  theme(
  plot.title = element_text(color = "black", size = 16, face = "bold",hjust = 0.5),
  axis.title = element_text(size = 16)
) 

plot3

ggsave("plot3.png", plot3)

```
High switching ratio --> low number samples 

Correlation between switch ratio and sampling total

```{r correlation coefficient}

cor.test(gain_dt$sampling_total,gain_dt$sampling_switch_ratio,  method = "pearson")

```

# Choice related properties 
## Descriptives 

```{r}
gain_dt[, .N , by = seen_exval_chosen]

gain_dt[, .N , by = seen_winner_chosen]

gain_dt[, .N , by = seen_roundwise_winner_chosen]

```

## How is the relationship between switching frequency and decision properties 

### Step 1 : Create sampling frequency with different decimals 

```{r sampling switch ratio with different decimals }
# create sampling_switch_ratio_3 --> same with the current sampling_switch_ratio
gain_dt[, sampling_switch_ratio_3 := sampling_switch_ratio]

# create sampling_switch_ratio_2, by round sampling_switch_ratio to 2 
gain_dt[, sampling_switch_ratio_2 := round(sampling_switch_ratio, 2)]

# create sampling_switch_ratio_1, by round sampling_switch_ratio to 1
gain_dt[, sampling_switch_ratio_1 := round(sampling_switch_ratio, 1)]
```

### Step 2 : calculate proportions 
```{r calculate proportions}
# Calculate proportion of seen_exval_chosen
gain_dt[, seen_exval_chosen_proportion_3 := round(sum(seen_exval_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_3]
gain_dt[, seen_exval_chosen_proportion_2 := round(sum(seen_exval_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_2]
gain_dt[, seen_exval_chosen_proportion_1 := round(sum(seen_exval_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_1]


# Calculate proportion of seen_winner_chosen
gain_dt[, seen_winner_chosen_proportion_3 := round(sum(seen_winner_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_3]
gain_dt[, seen_winner_chosen_proportion_2 := round(sum(seen_winner_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_2]
gain_dt[, seen_winner_chosen_proportion_1 := round(sum(seen_winner_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_1]


# Calculate proportion of seen_roundwise_winner_chosen
gain_dt[, seen_roundwise_winner_chosen_proportion_3 := round(sum(seen_roundwise_winner_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_3]
gain_dt[, seen_roundwise_winner_chosen_proportion_2 := round(sum(seen_roundwise_winner_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_2]
gain_dt[, seen_roundwise_winner_chosen_proportion_1 := round(sum(seen_roundwise_winner_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_1]

# Calculate proportion of seen_roundwise_winner_chosen
gain_dt[, seen_roundwise_winner_chosen_proportion_3 := round(sum(seen_roundwise_winner_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_3]
gain_dt[, seen_roundwise_winner_chosen_proportion_2 := round(sum(seen_roundwise_winner_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_2]
gain_dt[, seen_roundwise_winner_chosen_proportion_1 := round(sum(seen_roundwise_winner_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_1]

# Calculate proportion of seen_roundwise_winner_chosen
gain_dt[, exval_chosen := ifelse(choice == exval, TRUE, FALSE)]

gain_dt[, exval_chosen_proportion_3 := round(sum(exval_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_3]

gain_dt[, exval_chosen_proportion_2 := round(sum(exval_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_2]

gain_dt[, exval_chosen_proportion_1 := round(sum(exval_chosen == TRUE, na.rm = TRUE) / .N ,3), 
        by = sampling_switch_ratio_1]



```

### Step 3 : prepare graphs 


```{r}
# Create graphs: sampling_switch_ratio vs seen_exval_chosen_proportion & seen_winner_chosen_proportion




plot4 <- ggplot(gain_dt, aes(x = sampling_switch_ratio_2, y = seen_exval_chosen_proportion_2)) + 
  geom_point() + geom_smooth(method = lm, se = FALSE, color = "blue") + labs(title = "Summary strategy") + 
  ylab("Maximization of sampled average") +  xlab("Switching frequency") + theme_classic() + theme(
  plot.title = element_text(color = "black", size = 16, face = "bold",hjust = 0.5),
  axis.title = element_text(size = 14)
)+coord_fixed(ratio = 1) 

plot5 <- ggplot(gain_dt, aes(x = sampling_switch_ratio_2, y = exval_chosen_proportion_2)) + 
  geom_point() + geom_smooth(method = lm, se = FALSE, color = "blue") + labs(title = "Expected value") + 
  ylab("Maximization of expected average") +  xlab("Switching frequency")+ theme_classic() + theme(
  plot.title = element_text(color = "black", size = 16, face = "bold",hjust = 0.5),
  axis.title = element_text(size = 14)
)+coord_fixed(ratio = 1)


plot6 <- ggplot(gain_dt, aes(x = sampling_switch_ratio_2, y = seen_winner_chosen_proportion_2)) + 
  geom_point() + geom_smooth(method = lm, se = FALSE, color = "blue") + labs(title = "Round-wise strategy") + 
  ylab("Maximization of frequent winner") +  xlab("Switching frequency")+ theme_classic() + theme(
  plot.title = element_text(color = "black", size = 16, face = "bold",hjust = 0.5),
  axis.title = element_text(size = 14)
)+coord_fixed(ratio = 1)

plot4
plot5
plot6

ggsave("plot4.png", plot4,  width = 6, height = 6, units = "in")
ggsave("plot5.png", plot5,  width = 6, height = 6, units = "in")
ggsave("plot6.png", plot6,  width = 6, height = 6, units = "in")

```

# Proportion of true and false
```{r}
# Calculate proportions

gain_dt[, true_seen_exval_chosen_proportion := round(sum(seen_exval_chosen == TRUE & I1 == TRUE, na.rm = TRUE) / sum(seen_exval_chosen == TRUE, na.rm = TRUE) ,3), 
        by = sampling_switch_ratio_2]

gain_dt[, false_seen_exval_chosen_proportion := round(sum(seen_exval_chosen == TRUE & I1 == FALSE, na.rm = TRUE) / sum(seen_exval_chosen == TRUE, na.rm = TRUE) ,3), 
        by = sampling_switch_ratio_2]

gain_dt[, true_seen_winner_chosen_proportion := round(sum(seen_winner_chosen == TRUE & I1 == TRUE, na.rm = TRUE) / sum(seen_winner_chosen == TRUE, na.rm = TRUE) ,3), 
        by = sampling_switch_ratio_2]

gain_dt[, false_seen_winner_chosen_proportion := round(sum(seen_winner_chosen == TRUE & I1 == FALSE, na.rm = TRUE) / sum(seen_winner_chosen == TRUE, na.rm = TRUE) ,3), 
        by = sampling_switch_ratio_2]

gain_dt[, true_exval_chosen_proportion := round(sum(exval_chosen == TRUE & I1 == TRUE, na.rm = TRUE) / sum(exval_chosen == TRUE, na.rm = TRUE) ,3), 
        by = sampling_switch_ratio_2]

gain_dt[, false_exval_chosen_proportion := round(sum(exval_chosen == TRUE & I1 == FALSE, na.rm = TRUE) / sum(exval_chosen == TRUE, na.rm = TRUE) ,3), 
        by = sampling_switch_ratio_2]

```

```{r}
plot7 <- ggplot(gain_dt, aes(x = sampling_switch_ratio_2)) +
  geom_point(aes(y = true_seen_exval_chosen_proportion, color = "True"), size = 2) +
  geom_point(aes(y = false_seen_exval_chosen_proportion, color = "False"), size = 2) +
  geom_smooth(aes(y = true_seen_exval_chosen_proportion, color = "True"), method = "lm", se = FALSE) +
  geom_smooth(aes(y = false_seen_exval_chosen_proportion, color = "False"), method = "lm", se = FALSE) +
  scale_color_manual(name = "Indicator", 
                     values = c("True" = "#666666", "False" = "blue"), 
                     labels = c("True", "False")) +
  theme_classic() +
  labs(title = "Summary strategy", y = "Maximization of sampled average", x = "Switching frequency") + 
  theme(
    plot.title = element_text(color = "black", size = 16, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 16)
  ) +
  coord_fixed(ratio = 1)

plot7
#plot7
#ggsave("plot7.png", plot7,  width = 6, height = 6, units = "in")
```

```{r}
plot8 <- ggplot(gain_dt, aes(x = sampling_switch_ratio_2)) +
  geom_point(aes(y = true_seen_winner_chosen_proportion, color = "True"), color = "#666666", size = 2) +
  geom_point(aes(y = false_seen_winner_chosen_proportion, color = "False"), color = "blue", size = 2) +
  geom_smooth(aes(y = true_seen_winner_chosen_proportion, color = "True"), method = "lm", na.rm = TRUE, se = FALSE, color = "#666666") +
  geom_smooth(aes(y = false_seen_winner_chosen_proportion, color = "False"), method = "lm",na.rm = TRUE, se = FALSE, color = "blue") +
  theme_classic() +
  labs(title = "Round-wise strategy") + 
  ylab("Maximization of frequent winner") +  xlab("Switching frequency")+ theme_classic() + theme(
  plot.title = element_text(color = "black", size = 16, face = "bold",hjust = 0.5),
  axis.title = element_text(size = 16)
) +coord_fixed(ratio = 1) 

plot8
ggsave("plot8.png", plot8,  width = 6, height = 6, units = "in")
```

```{r}

plot9 <- ggplot(gain_dt, aes(x = sampling_switch_ratio_2)) +
  geom_point(aes(y = true_exval_chosen_proportion, color = "True"), size = 2, color = "#666666") +
  geom_point(aes(y = false_exval_chosen_proportion, color = "False"), size = 2, color = "blue") +
  geom_smooth(aes(y = true_exval_chosen_proportion, color = "True"), method = "lm", se = FALSE, color = "#666666") +
  geom_smooth(aes(y = false_exval_chosen_proportion, color = "False"), method = "lm", se = FALSE, color = "blue") +
  theme_classic() +labs(title = "Expected value") + 
  ylab("Maximization of expected average") +  xlab("Switching frequency")+ theme_classic() + theme(
  plot.title = element_text(color = "black", size = 16, face = "bold",hjust = 0.5),
  axis.title = element_text(size = 16)
) + coord_fixed(ratio = 1) 
plot9
ggsave("plot9.png", plot9,  width = 6, height = 6, units = "in")
```



# Data is for modelling
```{r}

write.csv(gain_dt, file = here("Data", "model_ready_data.csv"), row.names = FALSE)

```
