---
title: "Models_Final"
author: "Emine Cevik"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
here::i_am("Code/Models_Final.Rmd")
```

```{r setup, include = TRUE, echo = FALSE}
knitr::opts_chunk$set(message = FALSE)
```

```{r, warning = FALSE}
library(tidyverse)
library(openxlsx)
library(data.table)
library(here)
library(lme4)
library(stargazer)
```

# Read data
Read main data 
```{r, warning = FALSE, echo = TRUE}
#read the file 
model_dt <- as.data.table(read.csv(here("Data", "model_ready_data.csv")))
colnames(model_dt)

```

```{r}
model_dt[, I5 := ifelse(exval == seen_winner, 1, 0)]
```

# Logistic regression 
## Seen_exval
### Model1.1: 
DV: seen_exval_chosen, IV: sampling_switch_ratio

```{r}
log1.1 <- glm( seen_exval_chosen ~ sampling_switch_ratio , 
               data = model_dt, family = binomial)
summary(log1.1)
```
Results:

  - Coefficient of switching_ratio is significant and positive. It means that
  when switching ratio increases, the odds of seen_exval_chosen increases 
  - High switching ratio --> high chance of seen_exval_chosen 
  - High switching ratio correlates with low number samples. Does it mean that 
  the participants can also recognize seen_exval when they are frequent switcher?
    - It could be the case if seen_winner and seen_exval are same (I1)?

### Model1.2: 
DV: seen_exval_chosen, IV: sampling_switch_ratio + I1 (seen_exval = seen_winner)
```{r}
log1.2 <- glm( seen_exval_chosen ~ sampling_switch_ratio + I1 , 
               data = model_dt, family = binomial)
summary(log1.2)
```
Results:

  - When we control for I1, the effect of sampling_switch_ratio decreased
  - But still increasing switch ratio increases the odds of seen_exval_chosen
  - When seen_exval and seen_winner is same I1 is 1. So when both of them are same 
  the odds of seen_exval_chosen increases
  
  
### Model1.3: 
DV: seen_exval_chosen, IV: sampling_switch_ratio + I1 + I1*sampling_switch_ratio
```{r}
log1.3 <- glm( seen_exval_chosen ~ sampling_switch_ratio + I1 +
                 I1*sampling_switch_ratio, data = model_dt, family = binomial)
summary(log1.3)
```

Result:

When I1 = 0, so seen_exal and seen_winner is not same, increase in sampling_switch_ratio
decreases the odd of seen_exval_chosen 

When I1 = 1, it increases the odds of seen_exval_chosen but sampling_switch_ratio will 
also have a postive effect on this 

## Seen_winner
### Model2.1: 
DV: seen_winner_chosen, IV: sampling_switch_ratio

```{r}
log2.1 <- glm( seen_winner_chosen ~ sampling_switch_ratio , data = model_dt, family = binomial)
summary(log2.1)
```

### Model2.2: 
DV: seen_winner_chosen, IV: sampling_switch_ratio + I1 (seen_exval = seen_winner)
```{r}
log2.2 <- glm( seen_winner_chosen ~ sampling_switch_ratio + I1 , 
               data = model_dt, family = binomial)
summary(log2.2)
```

### Model2.3: 
DV: seen_winner_chosen, IV: sampling_switch_ratio + I1 + I1*sampling_switch_ratio
```{r}
log2.3 <- glm( seen_winner_chosen ~ sampling_switch_ratio + I1 +
                 I1*sampling_switch_ratio, data = model_dt, family = binomial)
summary(log2.3)
```

## Seen_roundwise_winnner
### Model3.1: 

DV: seen_roundwise_winner_chosen, IV: sampling_switch_ratio

```{r}
log3.1 <- glm( seen_roundwise_winner_chosen ~ sampling_switch_ratio , data = model_dt, family = binomial)
summary(log3.1)
```

### Model3.2: 
DV: seen_roundwise_winner_chosen, IV: sampling_switch_ratio + I2 (seen_exval = seen_roundwise_winner_chosen)
```{r}
log3.2 <- glm( seen_roundwise_winner_chosen ~ sampling_switch_ratio + I2 , 
               data = model_dt, family = binomial)
summary(log3.2)
```

### Model3.3: 
DV: seen_roundwise_winner_chosen, IV: sampling_switch_ratio + I2 + I2*sampling_switch_ratio
```{r}
log3.3 <- glm( seen_roundwise_winner_chosen ~ sampling_switch_ratio + I2 +
                 I2*sampling_switch_ratio, data = model_dt, family = binomial)
summary(log3.3)
```

## exval_chosen
### Model4.1: 

DV: exval_chosen, IV: sampling_switch_ratio

```{r}
log4.1 <- glm( exval_chosen ~ sampling_switch_ratio , data = model_dt, family = binomial)
summary(log4.1)
```

### Model4.2: 
DV: exval_chosen, IV: sampling_switch_ratio + I4 (seen_exval = exval)
```{r}
model_dt[, I4 := ifelse(exval == seen_exval, 1, 0)]

log4.2 <- glm( exval_chosen ~ sampling_switch_ratio + I4 , 
               data = model_dt, family = binomial)
summary(log4.2)
```

### Model4.3: 
DV: exval_chosen, IV: sampling_switch_ratio + I4 + I4*sampling_switch_ratio
```{r}
log4.3 <- glm( exval_chosen ~ sampling_switch_ratio + I4 +
                 I4 * sampling_switch_ratio, data = model_dt, family = binomial)
summary(log4.3)
```


### Model5.1: 

DV: exval_chosen, IV: sampling_switch_ratio
```{r}
log5.1 <- glm( exval_chosen ~ sampling_switch_ratio , data = model_dt, family = binomial)
summary(log5.1)
```

### Model5.2: 
DV: exval_chosen, IV: sampling_switch_ratio + I5 (seen_exval = exval)
```{r}

log5.2 <- glm( exval_chosen ~ sampling_switch_ratio + I5 , 
               data = model_dt, family = binomial)
summary(log5.2)
```

### Model5.3: 
DV: exval_chosen, IV: sampling_switch_ratio + I5 + I5*sampling_switch_ratio
```{r}
log5.3 <- glm( exval_chosen ~ sampling_switch_ratio + I5 +
                 I5 * sampling_switch_ratio, data = model_dt, family = binomial)
summary(log5.3)
```




# Analysis presented on the paper 
```{r}
model_dt[, I:= I1]
model_dt[, sampled_ev_chosen :=  seen_exval_chosen]
model_dt[, stated_ev_chosen :=  exval_chosen]
model_dt[, roundwise_chosen :=  seen_winner_chosen]
model_dt[, switching_frequency :=  sampling_switch_ratio]


model_sampled_ev <- model_dt[!is.na(I) & !is.na(sampled_ev_chosen), ]
model_stated_ev <- model_dt[ !is.na(I) & !is.na(stated_ev_chosen), ]
model_roundwise_chosen <- model_dt[!is.na(I) & !is.na(seen_winner_chosen), ]

m_1 <- glmer(sampled_ev_chosen ~ switching_frequency + I + I*switching_frequency + (1 | study) + (1 |study:participant),
                            data = model_sampled_ev, family = binomial)

m_2 <- glmer(stated_ev_chosen ~ switching_frequency + I + I*switching_frequency +  (1 | study) + (1 |study:participant),
                            data = model_stated_ev, family = binomial)

m_3 <- glmer(roundwise_chosen ~ switching_frequency + I + I*switching_frequency + (1 | study) + (1 |study:participant),
                            data = model_roundwise_chosen, family = binomial)

m_01 <- glmer(seen_exval_chosen ~  (1 | study) + (1 | study:participant),
                            data = model_sampled_ev, family = binomial)

m_02 <- glmer(seen_exval_chosen ~  (1 | study) + (1 | study:participant),
                            data = model_stated_ev, family = binomial)

m_03 <- glmer(seen_exval_chosen ~  (1 | study) + (1 | study:participant),
                            data = model_roundwise_chosen, family = binomial)
model_0 <- glmer(seen_exval_chosen ~  (1 | study) + (1 | study:participant),
                            data = model_stated_ev, family = binomial)

```

```{r}
anova(m_01,m_1)
```


```{r}

stargazer(m_1, m_2, m_3, type="latex",
          out="star_linear_3.doc")

```



